---
title: "Compare models and historical data"
output: html_document
---
```{r setup, message = FALSE, warning = FALSE, include = TRUE, echo = FALSE}
library(tidyverse)
library(here)
devtools::load_all()
```

## Load the historical data for most recent years
Let's see how the gravity and radiation model predictions fared for the three most recent years we have on hand: 2015, 2016, and 2017. 
We'll first load all the data and create matrices from the historical observations so it'll be easy to compute the residuals for the gravity and radiation models, which are already in matrix form.

Note: we'll be ignoring migrations within states, meaning the diagonals of these matrices will be zeros.

```{r}
# Load the data files

hist_2015 <- read.csv(here("results", "data_cleaning_results", "clean", "US_1516.csv"))
grav_2015 <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2015.csv")), nrow = 51, ncol = 51)
rad_2015 <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2015.csv")), nrow = 51, ncol = 51)

hist_2016 <- read.csv(here("results", "data_cleaning_results", "clean", "US_1617.csv"))
grav_2016 <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2016.csv")), nrow = 51, ncol = 51)
rad_2016 <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2016.csv")), nrow = 51, ncol = 51)

hist_2017 <- read.csv(here("results", "data_cleaning_results", "clean", "US_1718.csv"))
grav_2017 <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2017.csv")), nrow = 51, ncol = 51)
rad_2017 <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2017.csv")), nrow = 51, ncol = 51)

# Create matrices from historical data
hist_matrix_2015 <- create_historical_matrix(hist_2015)
hist_matrix_2016 <- create_historical_matrix(hist_2016)
hist_matrix_2017 <- create_historical_matrix(hist_2017)
```

### 2015 Data Comparisons
Now let's see what the heatmaps look like for each of these year's predictions.
```{r}
gravity_heatmap(grav_2015, hist_matrix_2015, "2015")
```

```{r}
radiation_heatmap(rad_2015, hist_matrix_2015, "2015")
```

### 2016 Data Comparisons
```{r}
gravity_heatmap(grav_2016, hist_matrix_2016, "2016")
```

```{r}
radiation_heatmap(rad_2016, hist_matrix_2016, "2016")
```

### 2017 Data Comparisons
```{r}
gravity_heatmap(grav_2017, hist_matrix_2017, "2017")
```

```{r}
radiation_heatmap(rad_2017, hist_matrix_2017, "2017")
```


From these maps, we can see that both models overestimate the number of migrants pretty often, as shown by the darker colors on the heatmaps. Ideally, both maps would be pretty consistently the lighter color all across the board, meaning the predictions are closer to the actual observed value. Overall, we can see that both do an okay job of estimating the number of migrants (about the same proportions of light to dark colors).

## Appendix: remaining years of data comparisons

### 2014
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_1415.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2014.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2014.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2014")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2014")
```

### 2013
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_1314.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2013.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2013.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2013")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2013")
```

### 2012
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_1213.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2012.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2012.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2012")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2012")
```

### 2011
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_1112.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2011.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2011.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2011")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2011")
```

### 2010
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_1011.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2010.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2010.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2010")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2010")
```

### 2009
```{r}
hist <- read.csv(here("results", "data_cleaning_results", "clean", "US_0910.csv"))
grav <- as.matrix(read.csv(here("results", "gravity_model_results", "alpha_0.3_year_2009.csv")), nrow = 51, ncol = 51)
rad <- as.matrix(read.csv(here("results", "radiation_model_results", "alpha_0.3_year_2009.csv")), nrow = 51, ncol = 51)
hist_matrix <- create_historical_matrix(hist)
gravity_heatmap(grav, hist_matrix, "2009")

```

```{r}
radiation_heatmap(rad, hist_matrix, "2009")
```
